Consumption and resistance across states and HRRs

### 2017-03-01 states/resistance/state-report.Rmd
simplifying

gram_negs = c('Escherichia coli', 'Klebsiella pneumoniae',
              'Proteus mirabilis', 'Enterobacter cloacae',
              'Klebisella oxytoca', 'Enterococcus faecium')
gram_poss = c('Coagulase-negative staphylococci',
              'Methicillin-resistant S. aureus (MRSA)',
              'Methicillin-susceptible S. aureus (MSSA)',
              'Streptococcus pneumoniae (non-meningitis)',
              'Staphylococcus aureus')
organisms = c(gram_negs, gram_poss)

consumption_drugs=c('trimethoprim/sulfamethoxazole', 'ciprofloxacin',
                    'levofloxacin', 'ceftriaxone', 'nitrofurantoin',
                    'amoxicillin', 'penicillin', 'doxycycline',
                    'azithromycin')
resistance_drugs=c('trimethoprim-sulfamethoxazole', 'ciprofloxacin',
                   'levofloxacin', 'ceftriaxone', 'nitrofurantoin',
                   'ampicillin', 'penicillin',
                   'tetracycline/doxycycline', 'erythromycin')

consumption_resistance = data_frame(consumption_drug=consumption_drugs,
                                    resistance_drug=resistance_drugs)

# get all bug/drug combos (i.e., all bug/cons drug combos with appropriate res drugs)
bug_drug = expand.grid(bug=organisms,
                       consumption_drug=consumption_drugs,
                       stringsAsFactors=FALSE) %>%
  as_data_frame() %>%
  left_join(consumption_resistance, by='consumption_drug') %>%
  select(bug, consumption_drug, resistance_drug)
```

@@

I performed this matching by hand.

`r consumption_resistance %>% kable`

```{r load_data}
coherence = read_tsv('../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
  rename(resistance_drug=drug) %>%
  mutate(resistance_drug=tolower(resistance_drug),
         state=level %in% c('state', 'hrr', 'zipcode'),
         hrr=level %in% c('hrr', 'zipcode')) %>%
  select(bug, resistance_drug, state, hrr) %>%
  gather('place', 'coherent', state, hrr)

@@

# Antibiogram data, with individual records
abg = read_tsv('../../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
  rename(resistance_drug=drug) %>%
  left_join(hrr, by='zipcode') %>%
  select(bug, resistance_drug, state, hrr, percent_nonsusceptible)

@@

# State-level consumption data
state_consumption = read_tsv('../bene-consumption/state_usage_2011.tsv') %>%
  rename(consumption_drug=antibiotic) %>%
  select(state, consumption_drug, cpkp, did)

hrr_consumption = read_tsv('../bene-consumption/hrr_usage_2011.tsv') %>%
  rename(consumption_drug=antibiotic) %>%
  select(hrr, consumption_drug, cpkp, did)

state_dat = bug_drug %>%
  left_join(abg, by=c('bug', 'resistance_drug')) %>%
  left_join(state_consumption, by=c('state', 'consumption_drug'))

hrr_dat = bug_drug %>%
  left_join(abg, by=c('bug', 'resistance_drug')) %>%
  left_join(hrr_consumption, by=c('hrr', 'consumption_drug'))
```

@@

- *S. aureus* and azithromycin
- Potentially, a backward relationship for *S. aureus* and TMP-SMX
- *S. pneumoniae* and azithromycin
- *E. coli* and quinolones

Interestingly, this accords with some of the conclusions from @bell-systematic-2014, namely:

>>>>

gram_negs = c('Escherichia coli', 'Klebsiella pneumoniae',
              'Proteus mirabilis', 'Enterobacter cloacae',
              'Klebisella oxytoca', 'Enterococcus faecium')
gram_poss = c('Coagulase-negative staphylococci',
              'Methicillin-resistant S. aureus (MRSA)',
              'Methicillin-susceptible S. aureus (MSSA)',
              'Streptococcus pneumoniae (non-meningitis)',
              'Staphylococcus aureus')
organisms = c(gram_negs, gram_poss)
```

## Organisms

I selected a list of organisms from the ones that were most abundant in the data:

`r data_frame(organsisms=organisms) %>% kable`

@@

I performed this matching by hand. I placed the consumption drugs into groups:

`r consumption_groups %>% kable`

And did the same with the resistance drugs:

`r resistance_groups %>% kable`

I chose not to merge the penicillin and amoxicillin groups because antibiograms often reported very different susceptibilities for the two drugs.

## Consumption and resistance data

```{r load_data}
consumption_groups = read_tsv('consumption_groups.tsv')
resistance_groups = read_tsv('resistance_groups.tsv')

coherence = read_tsv('../../../../antibiogram/analysis/geographic-coherence/coherences.tsv') %>%
  rename(resistance_drug=drug) %>%
  mutate(resistance_drug=tolower(resistance_drug),
         state=level %in% c('state', 'hrr', 'zipcode'),
         hrr=level %in% c('hrr', 'zipcode')) %>%
  select(bug, resistance_drug, state, hrr) %>%
  gather('place', 'coherent', state, hrr)

@@

# Antibiogram data, with individual records
abg = read_tsv('../../../../antibiogram/data/abg.tsv', col_types=cols(zipcode='c')) %>%
  mutate(drug=tolower(drug), percent_nonsusceptible=100-percent_susceptible) %>%
  left_join(hrr, by='zipcode') %>%
  right_join(resistance_groups, by='drug') %>%
  filter(big %in% organisms) %>%
  select(bug, drug_group, state, hrr, percent_nonsusceptible)

@@

# State-level consumption data
state_consumption = read_tsv('../bene-consumption/state_usage_2011.tsv') %>%
  filter(!is.na(state)) %>%
  right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
  group_by(state, drug_group) %>%
  summarize_at(vars(cpkp, did), sum)

hrr_consumption = read_tsv('../bene-consumption/hrr_usage_2011.tsv') %>%
  filter(!is.na(hrr)) %>%
  right_join(consumption_groups, by=c('antibiotic'='drug')) %>%
  group_by(hrr, drug_group) %>%
  summarize_at(vars(cpkp, did), sum)

state_dat = left_join(abg, state_consumption, by=c('drug_group', 'state'))
hrr_dat = left_join(abg, hrr_consumption, by=c('drug_group', 'hrr'))
```

## Coherence of bug/drug combinations in resistance data

I want to be able to filter the bug/drug combinations by whether they are coherent in the resistance data. Does the geography (state or HRR) actually have any effect on the

@@

- *S. aureus* and azithromycin
- Potentially, a backward relationship for *S. aureus* and TMP-SMX
- *S. pneumoniae* and azithromycin
- *E. coli* and quinolones
Interestingly, this accords with some of the conclusions from @bell-systematic-2014, namely:
